### 2014-06-02 patchy-selection-paper.tex
added shared lineage

%%%%%%%
\subsection{Establishment of a locally adaptive allele due to migrational influx}
\label{ss:patchymigration}
\gc{There's a bunch of switching between x and r in this section}
Now suppose that there are two patches, each of area $A$ and separated by distance $R$. 
If the $B$ allele has arisen and become established in the first patch, but has not yet appeared in the second,
we would like to know the time scale on which copies of $B$ move between patches by migration
(so, ignoring the possibility that a $B$ allele arises independently by mutation in the first patch).
To determine this, we dissect the process by which an allele transits between the patches,
in the process obtaining other useful information about the genealogy of B alleles.

\gc{Could you make the CI or sample path in Figure 1B be different colors, the grey and black are too close to each other.}
\plr{how's that?  I'm not sure what to do about the grey; the real problem is that the CIs are too small. If I ran them for less time they would be\dots}

@@

This suggests that families can be well-described by motion of a central ``trunk'',
dies at a random time whose distribution is given by $k_e(t)$,
and conditional on survival of this trunk, the family is composed of side ``branches'' off of this trunk,
as depicted in Figure~\ref{fig:branching_decomp}.
We can understand this quite concretely, by a method described rigorously by \citet{geiger1999elementary} \citep[see also][]{chauvin1991growing}.
When we condition on survival until $t$, we condition on existence of
at least one ``trunk'' lineage surviving from time $0$ to time $t$ (the long red
lineage in Figure \ref{fig:branching_decomp}).
Given this lineage, nothing distinguishes any of the other reproduction events --
each individual along the trunk lineage must give birth to at least one offspring,
and all other reproduction events are unconditioned.
In other words, the genealogy of a family that survives until $t$
can be decomposed into a ``trunk'' that lasts until $t$
which sprouts independent ``branches'' that may or may not survive until $t$.
Concretely, if we write $Z_t$ for the number of offspring in the branching process alive at time $t$,
% (so $\P\{Z_t = 0\} = k_e(t)$),
then this gives the following decomposition:
\begin{align}
  \left( Z_t \; \vert \; Z_t>0 \right) \deq 1 + \sum_{s=0}^t \sum_{k=1}^{W_s-1} Z^{(s,k)}_{t-s},
\end{align}
where each $Z^{(s,k)}$ is an independent copy of $Z$ 
that arose from the main branch in generation $s$,
and each $W_s$ is an independent copy of the offspring distribution, conditioned on $W_s \ge 1$.
\gc{Is there no 1+ in the equation above for the main branch? } \plr{oops; yep.}

@@

First consider equation \eqref{eqn:gestalt_q}.
Suppose that $Z$ is one such spatial branching process as above, started at time 0 with a single individual at $x$,
defined so that $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the spatial region $A$.
The \emph{mean occupation density} of $Z$ can be thought of 
as the expected total number of offspring of a family beginning at $x$ that ever live at $y$.
Concretely, it is the density of $\E[\int_0^t Z_t dt]$,
i.e.\ the smooth function $u(x,y)$ that, integrated over a set $A$, 
gives the expected total number of individuals ever living in $A$:
\begin{align} \label{eqn:occupation_density_defn}
    \E\left[ \int_0^\infty Z_t(A) dt \right] = \int_A u(x,y) dy .
\end{align}
\gc{do we need to talk about u a little more it }
\plr{not sure what to say}
Since reproduction and spatial motion are independent,
$u(x,y)$ is the sum across all generations of the expected number of individuals alive in that generation,
multiplied by the probability density that spatial motion takes a lineage from $x$ to $y$,
i.e.\
\begin{align}
    u(x,y) = \int_0^t e^{-s_m t} p_t(x,y) dt .
\end{align}
The equilibrium frequency as decomposed in expression \eqref{eqn:gestalt_q} 
is the sum of mean occupation densities of a constant outflux of branching processes
from $\partial B(r_0)$:
\begin{align} \label{eqn:occupation_integral}
    q(y) &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  . % \\
         % &= \int_{\partial B(r_0)} \gamma(x) \int_0^t e^{-s_m t} p_t(x,y) dt dx .
\end{align}
\gc{define $\gamma(x)$ presumably the outflux of branches leaving ball}
\plr{it was defined at the start of this paragraph.}

@@

\begin{figure}[ht]
  \begin{center}
    \includegraphics{sim-transit.pdf}
  \end{center}
  \caption{
  BLAH BLAH WIGGLY LINE
  \label{fig:lineagesmotion}
  }
\end{figure}

>>>>

%%%%%%%
\subsection{Establishment of a locally adaptive allele due to migrational influx}
\label{ss:patchymigration}

Now suppose that there are two patches, each of area $A$ and separated by distance $R$. 
If the $B$ allele has arisen and become established in the first patch, but has not yet appeared in the second,
we would like to know the time scale on which copies of $B$ move between patches by migration
(so, ignoring the possibility that a $B$ allele arises independently by mutation in the first patch).
To determine this, we dissect the process by which an allele transits between the patches,
in the process obtaining other useful information about the genealogy of B alleles.

@@

This suggests that families can be well-described by motion of a central ``trunk'',
dies at a random time whose distribution is given by $k_e(t)$,
and conditional on survival of this trunk, the family is composed of side ``branches'' off of this trunk,
as depicted in Figure~\ref{fig:branching_decomp}.
We can understand this quite concretely, by a method described rigorously by \citet{geiger1999elementary} \citep[see also][]{chauvin1991growing}.
When we condition on survival until $t$, we condition on existence of
at least one ``trunk'' lineage surviving from time $0$ to time $t$ (the long red
lineage in Figure \ref{fig:branching_decomp}).
Given this lineage, nothing distinguishes any of the other reproduction events --
each individual along the trunk lineage must give birth to at least one offspring,
and all other reproduction events are unconditioned.
In other words, the genealogy of a family that survives until $t$
can be decomposed into a ``trunk'' that lasts until $t$
which sprouts independent ``branches'' that may or may not survive until $t$.
Concretely, if we write $Z_t$ for the number of offspring in the branching process alive at time $t$,
% (so $\P\{Z_t = 0\} = k_e(t)$),
then this gives the following decomposition:
\begin{align}
  \left( Z_t \; \vert \; Z_t>0 \right) \deq 1 + \sum_{s=0}^t \sum_{k=1}^{W_s-1} Z^{(s,k)}_{t-s},
\end{align}
where each $Z^{(s,k)}$ is an independent copy of $Z$ 
that arose from the main branch in generation $s$,
and each $W_s$ is an independent copy of the offspring distribution, conditioned on $W_s \ge 1$.

@@

First consider equation \eqref{eqn:gestalt_q}.
Suppose that $Z$ is one such spatial branching process as above, started at time 0 with a single individual at $x$,
defined so that $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the spatial region $A$.
The \emph{mean occupation density} of $Z$ can be thought of 
as the expected total number of offspring of a family beginning at $x$ that ever live at $y$.
Concretely, it is the density of $\E[\int_0^t Z_t dt]$,
i.e.\ the smooth function $u(x,y)$ that, integrated over a set $A$, 
gives the expected total number of individuals ever living in $A$:
\begin{align} \label{eqn:occupation_density_defn}
    \E\left[ \int_0^\infty Z_t(A) dt \right] = \int_A u(x,y) dy .
\end{align}
\gc{do we need to talk about u a little more it }
\plr{not sure what to say}
Since reproduction and spatial motion are independent,
$u(x,y)$ is the sum across all generations of the expected number of individuals alive in that generation,
multiplied by the probability density that spatial motion takes a lineage from $x$ to $y$,
i.e.\
\begin{align}
    u(x,y) = \int_0^t e^{-s_m t} p_t(x,y) dt .
\end{align}
The equilibrium frequency as decomposed in expression \eqref{eqn:gestalt_q} 
is the sum of mean occupation densities of a constant outflux of branching processes
from $\partial B(r_0)$:
\begin{align} \label{eqn:occupation_integral}
    q(y) &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  . % \\
         % &= \int_{\partial B(r_0)} \gamma(x) \int_0^t e^{-s_m t} p_t(x,y) dt dx .
\end{align}

@@

\begin{figure}[ht]
  \begin{center}
    \includegraphics{sim-transit.pdf}
  \end{center}
  \caption{
  BLAH BLAH WIGGLY LINE
  \plr{can we please keep that caption? ok, fine\dots}
  \textbf{Adaptation by migration:}
  An initially unadapted patch (10 demes at the left end of the range) is colonized around generation 5000;
  a representative lineage from the new patch is shown transiting between the two and coalescing with a lineage
  in the unadapted patch.
  As in Figure~\ref{fig:sim_snapshots}, color and contours show allele frequency across space and time;
  other simulation parameters are as in Figures~\ref{fig:sim_occupation_freqs}B and \ref{fig:sim_snapshots}.
  \label{fig:lineagesmotion}
  }
\end{figure}
